# NNYC Service Request Statistical Analysis

## Plotting

```{r}
library('ggplot2')
library('qqplotr')
library('dplyr')
library(scales)
data = read.csv('../CSV/NYPD_Complaints_Eric.csv')
```

```{r}
data
```

```{r}

op <- ggplot(data, aes(x = duration.in.hours)) +
  geom_histogram(
    bins = 30,                  # number of bins
    fill = "#1f77b4",           # professional blue color
    color = "white"             # white borders around bars
  ) +
  labs(
    title = "Distribution of Complaint Resolution Times with outliers",
    x = "Resolution Time (hours)",
    y = "Number of Complaints"
  ) +
  theme_minimal(base_size = 14) +  # clean theme with readable text
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 18),
    axis.title = element_text(face = "bold", size = 14),
    axis.text = element_text(size = 12)
  )


ggsave(
  filename = "../plots/op_hist_dist.png",
  plot = op,
  width = 12, height = 8,  # size in inches
  units = "in",
  dpi = 300
)
op
```

| Outliers were observed and excluded to mitigate their influence on statistical estimates.

```{r}
Q1 <- quantile(data$duration.in.hours, .25)
Q3 <- quantile(data$duration.in.hours, .75)
IQR <- IQR(data$duration.in.hours)
```

```{r}
  lower_bound <- Q1 - 1.5 * IQR
  upper_bound <- Q3 + 1.5 * IQR
```

```{r}
data <- data[data$duration.in.hours > lower_bound & data$duration.in.hours < upper_bound,]
data
```

```{r}
p <- ggplot(data, aes(x = duration.in.hours)) +
  geom_histogram(
    bins = 30,                  # number of bins
    fill = "#1f77b4",           # professional blue color
    color = "white"             # white borders around bars
  ) +
  labs(
    title = "Distribution of Complaint Resolution Times",
    x = "Resolution Time (hours)",
    y = "Number of Complaints"
  ) +
  theme_minimal(base_size = 14) +  # clean theme with readable text
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 18),
    axis.title = element_text(face = "bold", size = 14),
    axis.text = element_text(size = 12)
  )

# Export to PDF for poster
ggsave(
  filename = "../plots/hist_dist.png",
  plot = p,
  width = 12, height = 8,  # size in inches
  units = "in",
  dpi = 300
)
p
```

```{r}
sample_data = sample(data$duration.in.hours, 1000, replace = FALSE, prob = NULL)
```

```{r}
shapiro_test <- shapiro.test(sample_data)
shapiro_test
```

We reject the null hypothesis that the data is normally distributed

```{r}
ggplot(data, aes(x = duration.in.hours)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "white") +
  facet_wrap(~ borough) +
  labs(
    x = "Duration (hours)",
    y = "Count",
    title = "Distribution of Duration by Borough"
  ) +
  theme_minimal()
```

```{r}
N_comp<-data %>%
  count(borough) %>%
  ggplot(aes(x = reorder(borough, n), y = n)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = comma(n)),
            hjust = -0.1,
            size = 4) +
  coord_flip() +
  scale_y_continuous(labels = comma,
                     expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Number of NYPD Complaints by Borough",
    x = "Borough",
    y = "Number of Complaints"
  ) +
  theme_minimal(base_size = 14)
ggsave(
  filename = "../plots/n_complaint.png",
  plot = N_comp,
  width = 12, height = 8,  # size in inches
  units = "in",
  dpi = 300
)
N_comp
```

## Answering Questions

### Are the boroughs independent of the complaint type?

```{r}

top5 <- data %>%
  count(complaint_type, sort = TRUE) %>%
  slice_head(n = 5) %>%
  pull(complaint_type)

# Step 2: Group remaining complaint types as "Other"
df_plot <- data %>%
  mutate(complaint_group = ifelse(
    complaint_type %in% top5,
    complaint_type,
    "Other"
  ))

# Step 3: Define professional color palette
palette_colors <- c(
  "Noise - Residential" = "#1f77b4",        # blue
  "Illegal Parking" = "#ff7f0e",  # orange
  "Blocked Driveway" = "#9467bd", # green
  "Noise - Street/Sidewalk" = "#d62728", # red
  "Noise - Vehicle" = "#2ca02c",   # purple
  "Other" = "gray70"              # gray for everything else
)
# Step 4: Create proportional stacked bar chart
stacked_bar<-ggplot(df_plot, aes(x = borough, fill = complaint_group)) +
  geom_bar(position = "fill", color = "white", size = 0.3) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = palette_colors) +
  labs(
    title = "Borough-Level Differences in NYPD Complaint Types",
    x = "Borough",
    y = "Percentage of Complaints",
    fill = "Complaint Type"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, face = "bold"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.y = element_text(face = "bold"),
    legend.position = "right",
    axis.title = element_text(size = 20),
    axis.text  = element_text(size = 20),
    legend.title = element_text(size = 20),
    legend.text  = element_text(size = 16)
  )
ggsave(
  filename = "../plots/stacked_bar.png",
  plot = N_comp,
  width = 12, height = 8,  # size in inches
  units = "in",
  dpi = 300
)

stacked_bar
```

```{r}
complaint_by_borough <- table(data$complaint_type, data$borough)
complaint_by_borough
```

```{r}
chi_test <- chisq.test(x = complaint_by_borough)
chi_test
```

| There is a relationship between the type of crime and the borough

### Is the distribution of resolution time the same across all boroughs?

### Does the Borough have an affect on resolution?

```{r}
kurskal_test <- kruskal.test(duration.in.hours ~ borough, data = data)
kurskal_test
```

| There is a statistically significant difference in complaint resolution time across boroughs

```{r}
library(ggplot2)

cap <- quantile(data$duration.in.hours, 0.99, na.rm = TRUE)

data_plot <- data %>%
  mutate(
    duration_capped = pmin(duration.in.hours, cap)
  )
bp_duration<-ggplot(data_plot, aes(x = borough, y = duration_capped, fill = borough)) +
  geom_boxplot(outlier.shape = NA, width = 0.6) +
  stat_summary(
    fun = median,
    geom = "point",
    shape = 23,
    size = 3,
    fill = "white"
  ) +
  stat_boxplot(
    geom = "errorbar",
    width = 0.25,
    linewidth = 1.2
  )+
  labs(
    title = "Response Time Distribution by Borough",
    subtitle = "Median response time highlighted",
    x = "Borough",
    y = "Response Time (hours)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold")
  )

ggsave(
  filename = "../plots/bp_duration.png",
  plot = bp_duration,
  width = 12, height = 8,  # size in inches
  units = "in",
  dpi = 300
)
bp_duration
```

![]()

## Does the likelihood of police action differ by borough?

```{r}
tbl <- table(data$borough, data$took_action)
tbl
```

```{r}
chisq.test(tbl)
```

```{r}
tk_action<-data %>%
  group_by(borough) %>%
  summarise(pct_action = mean(took_action == 1)*100) %>%
  ggplot(aes(x = reorder(borough, pct_action), y = pct_action)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = paste0(round(pct_action,1), "%")), hjust = -0.1) +
  coord_flip() +
  labs(title = "Percentage of Complaints Where Police Took Action by Borough",
       x = "Borough", y = "Percentage") +
  ylim(0,100) +
  theme_minimal(base_size = 14)+
  theme_minimal(base_size = 18) +  # increases base font size for axes and ticks
  theme(
    plot.title = element_text(size = 24, face = "bold"),  # bigger title
    axis.title = element_text(size = 20, face = "bold"),   # bigger axis titles
    axis.text = element_text(size = 16)                    # bigger tick labels
  )

ggsave(
  filename = "../plots/tk_action.png",
  plot = tk_action,
  width = 15, height = 10,  # size in inches
  units = "in",
  dpi = 300
)

tk_action
```

## Is the difference in closing complaints a coincidence?

```{r}
 data_bill = read.csv("../CSV/NYPD_Complaints_Bill.csv")
```

```{r}
data_bill
```

```{r}
Q1 <- quantile(data_bill$duration.in.hours, .25)
Q3 <- quantile(data_bill$duration.in.hours, .75)
IQR <- IQR(data_bill$duration.in.hours)
lower_bound <- Q1 - 1.5 * IQR
upper_bound <- Q3 + 1.5 * IQR
```

```{r}
data_bill <- data_bill[data_bill$duration.in.hours > lower_bound & data_bill$duration.in.hours < upper_bound & data_bill$duration > 0,]
data_bill
```

```{r}
wilcox_test <- wilcox.test(data$duration.in.hours, data_bill$duration.in.hours )
wilcox_test
```

Different median under different Mayors. We can reject the idea that the difference is a coincidence

```{r}
median(data$duration.in.hours, na.rm = TRUE)
median(data_bill$duration.in.hours, na.rm = TRUE)
```

```{r}
df <- data.frame(
  duration = c(data$duration.in.hours,
               data_bill$duration.in.hours),
  person = factor(
    c(rep("Eric Adams", length(data$duration.in.hours)),
      rep("Bill de Blasio", length(data_bill$duration.in.hours)))
  )
)
```

```{r}
library(ggplot2)
library(dplyr)

y_max <- quantile(df$duration, 0.95)

b <- ggplot(df, aes(x = person, y = duration, fill = person)) +
  geom_boxplot(width = 0.5, alpha = 0.5, outlier.shape = NA) +
   
  stat_boxplot(geom = "errorbar", width = 0.25, linewidth = 1.2) +
  labs(
    title = "Service Request Duration by Mayoral Period",
    x = NULL,
    y = "Duration (hours)"
  ) +
  scale_fill_manual(values = c("Eric Adams" = "#4DAF4A", "Bill de Blasio" = "#377EB8")) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
    panel.grid.minor = element_blank(),
    axis.title = element_text(size = 20),
    axis.text  = element_text(size = 18)
  ) +
  ylim(0, y_max)

ggsave("../plots/box_plot_clean.png", plot = b, width = 10, height = 8, units = "in", dpi = 300)

b

```

## Do certain complaint types get resolved faster than others?

```{r}
df_median <- data %>%
  group_by(complaint_type) %>%
  summarise(median_duration = median(duration.in.hours),
            n = n()) %>%
  ungroup()

# Optional: keep only top 15 complaint types by count
top_types <- df_median %>% 
  top_n(15, n) %>% 
  pull(complaint_type)

df_median <- df_median %>%
  filter(complaint_type %in% top_types)
```

```{r}
library(ggplot2)

b <- ggplot(df_median, aes(
      x = reorder(complaint_type, median_duration),
      y = median_duration,
      fill = median_duration
    )) +
  geom_col() +
  geom_text(aes(label = round(median_duration,1)), 
            hjust = -0.1, size = 5) +  # show median on top of bar
  coord_flip() +  # horizontal bars
  scale_fill_gradient(low = "#4DAF4A", high = "#377EB8") +
  labs(
    title = "Median Duration of Complaints by Type",
    x = "Complaint Type",
    y = "Median Duration (hours)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 24, face = "bold"),
    axis.title = element_text(size = 20),
    axis.text = element_text(size = 16),
    legend.position = "none"
  ) +
  ylim(0, max(df_median$median_duration) * 1.1)  # add space for labels
```

```{r}
ggsave(
  filename = "../plots/median_duration_by_type.png",
  plot = b,
  width = 12,
  height = 8,
  units = "in",
  dpi = 300
)
```

```{r}
kruskal.test(duration.in.hours ~ complaint_type, data = data)
```
